### 2013-11-02 NeurounitFrontiers/Paper.tex
Addressed to do items

Software engineers validate software by developing simple executable tests, called ``unit tests'', that individually check a portion of a program against a single correctness criterion. Suites of such unit tests that collectively capture the desired program behavior provide evidence for its validity and correctness. Drawing inspiration from this practice, we develop the \textit{SciUnit} framework for developing suites of ``model validation tests'' -- 
executable functions, here written in Python, that compare model predictions against single empirical observations to  produce a score indicating agreement between the model and the data. Suites of such validation tests are used as a summary of modeling progress within a scientific community. 
Building on this framework, we develop \textit{NeuronUnit}, a library that supports validation testing for neurophysiology\todo{is this accurate?}.

@@

Today, scientists contribute a new model to the research community by submitting a paper describing how the model works, along with selected figures demonstrating its scope and validity and comparing its validity to other models with the same scope. Reviewers are then responsible for discovering relevant data and competing models that the paper did not adequately consider, drawing on their knowledge of prior publications. Unfortunately, in many areas, the number of publications being generated every year can overwhelm  even the most conscientious  scientists \todo{citation/stats?}.

@@

Professional software developers face similar issues. They must understand the scope of each component of a software project and validate it by measuring how well each component achieves its stated goals. But software developers do not validate components by simply choosing a few interesting inputs and presenting the outputs to reviewers. Rather, they typically follow a \emph{test-driven development} methodology by creating a suite of executable \emph{unit tests} that collectively specify the scope of each component and validate the functionality of its implementation as it is being developed and modified. Each test individually checks that a small portion of the program meets a single correctness criterion. For example, a unit test might verify that one function within the program correctly handles malformed inputs. Collectively, the test results serve as a summary of the validity of the project as it progresses through its development cycle. Developers can determine which features are unimplemented or buggy by examining the set of failed tests, and progress can be measured in terms of how many tests the program passes over time. This methodology is widely adopted in practice \cite{tdd}\todo{cite tdd}.

@@

Test-driven methodologies have started to see success in science as well. Modeling competitions in neuroscience, for example, are typically organized around a collection of simple validation criteria, implemented as executable tests. These competitions continue to drive important advances and improve scientists' understanding of the relative merits of different models. For example, 
the quantitative single neuron modeling competition (QSNMC)\cite{jolivet_quantitative_2008} investigates the complexity-accuracy tradeoff among reduced models of excitable membranes; 
the ``Hopfield'' challenge\cite{hopfield_what_2000} tested techniques for generating neuronal network form given its function; 
the Neural Prediction Challenge sought the best stimulus reconstructions, given neuronal activity\cite{neural_prediction_url}; the Diadem challenge is advancing the art of neurite reconstruction\cite{diadem_url}; and 
examples from other subfields of biology abound\cite{dream_url}.

% \begin{figure}%{r}{0.6\textwidth}
% %\vspace{-38px}
% \includegraphics[scale=0.75]{sciunit_overview.pdf}
% \caption{Proposal overview. 
% \small{i) The \textit{SciUnit} framework can generate discipline-specific libraries for tests generation and model interfacing.  
% \textit{NeuronUnit} is proposed and described here.  Experimental data guides test development; model generation informs and is informed by these \textit{SciUnit} libraries.  
% Each \text{SciUnit} test suite repository on GitHub is indexed by the \textit{SciDash} web application.  
% Test suites for OSB and QSNMC (described in the text) are being built. 
% \textit{SciDash} automatically runs the test collections in each suite and publicly displays the results.}}
% %\vspace{-10px}
% \label{fig:sciunit_overview}
% \end{figure}
% \leavevmode
% \vspace{-0px}
\begin{figure}
\vspace{-38px}
\centering
\includegraphics[scale=0.6]{diagram1.pdf}
\vspace{-55px}
\caption{Paper overview.}
\label{fig:sciunit_overview}
\end{figure}
\leavevmode
\vspace{-0px}

Each of these examples has leveraged \emph{ad hoc} infrastructure to support test generation. While the specific criteria used to evaluate models varies widely between disciplines in neuroscience, the underlying test-driven methodology has many common features that could be implemented once. Recognizing this, we have developed a discipline-agnostic framework for developing scientific validation test suites called \textit{SciUnit}\cite{sciunit_url}. To demonstrate the power of this framework and show how it can be used to bridge several existing neuroinformatics tools, we then  developed \textit{NeuronUnit\-}. This library allows scientists to build \textit{SciUnit} tests that validate \text{NeuroML}  models against electrophysiological data collected by the \textit{NeuroElectro Project} \cite{neuroelectro_url}, leveraging facilities from the \textit{NeuroTools} project to..\todo{what does NeuroTools do?}\todo{refer to and describe figure 1}\todo{add caption to figure 1}.% In addition to this open testing toolkit we also provide a library of tests for models of spiking single neurons.

@@

A simple {family} of models, \verb{LinearModel}, implementing the capability in Fig. \ref{fig:capability} as well as another capability shown in Fig. \ref{fig:training}, is shown in Fig. \ref{fig:simple_model}. 
Models in this family, \verbx{LinearModel}, produce a spike count by applying a linear transformation to the mean of the provided input current. 
The family is parameterized by a scale factor and offset, both scalars. 
To create a \emph{particular} model, the modeler may choose particular parameter values, just as with test families:
\begin{python}
CA1_linear_model_heuristic = LinearModel(3.0, 1.0)
\end{python}
Here, the parameters to the model were picked by the modeler heuristically, or based on externally-available knowledge. 
An alternative test design to the one in Fig. \ref{fig:rate_test} would add a second required capability (Fig. \ref{fig:training}) and a corresponding training phase to the \verb|run_test| method. 
This design would be relevant only for those models for which parameters can be adjusted without human involvement. 
Whether to build a training phase into the test protocol is a choice left to each modeler. 
Although Fig. \ref{fig:rate_test} does not include a training phase, if training data is available then models that nevertheless implement a training capability (like \verb|LinearModel|) can simply be trained explicitly by calling the capability method just like any other Python method:
\begin{python}
CA1_linear_model_fit = LinearModel()
CA1_linear_model_fit.train_with_currents(CA1_training_in, CA1_training_out)
\end{python}
\begin{figure}
\begin{python}
class LinearModel(sciunit.Model, SpikeCountFromCurrent, 
    TrainSpikeCountFromCurrent):
  def __init__(self, scale=None, offset=None): 
    self.scale, self.offset = scale, offset

@@

% Section 4.
\section{Discussion}
The validation criteria is made explicit in the specification of a test, so modelers need not guess which criteria are being used to validate or invalidate their model. 
Validation criteria are subject to debate (indeed, the QSNMC criteria changed between 2007 and 2008 due to such debates), and neuroscientists who wish to promote different criteria need only derive alternative tests. A community consensus will slowly emerge in the form a commonly-accepted test suite (see Sec. \ref{sec:scidash}). \todo{integrate this somewhere down here}

@@

\subsection{Arbitrary Scoring Criteria for Tests}
A test first assesses goodness-of-fit, and applies a normalization (e.g. pass/fail, 0.0-1.0) to generate a score. 
Arbitrary choices at both stages may benefit some models over others.  \textbf{First}, however, rank-ordering is constant across many goodness-of-fit metrics, meaning that choice of metric will rarely cause an otherwise passing model to fail and vice versa. 
For example, given a data mean and variance, ordering model output by Z-score or p-value will yield the same relative ranking of models. 
Indeed, rank ordering of models may prove more valuable than test scores themselves. 
\textbf{Second}, suite repositories are open (e.g. Fig. \ref{fig:scidash_repo}), so tests can be cloned and new statistical choices implemented. 
Statistics as applied to neuroscience have been notoriously ``fast and loose''; 
identification and correction of flawed methodology is becoming increasingly common\cite{button_power_2013,kriegeskorte_circular_2009,galbraith_study_2010,fish_fmri_url}, and is accelerated by an open framework. 
The community can judge which test version is most appropriate, i.e. what a model \textit{should} do -- this process documented via standard moderation techiniques used on GitHub -- and the \textit{SciUnit} framework determines whether the model \textit{does} it.

>>>>

Software engineers validate software by developing simple executable tests, called ``unit tests'', that individually check a portion of a program against a single correctness criterion. Suites of such unit tests that collectively capture the desired program behavior provide evidence for its validity and correctness. Drawing inspiration from this practice, we develop the \textit{SciUnit} framework for developing suites of ``model validation tests'' -- 
executable functions, here written in Python, that compare model predictions against single empirical observations to  produce a score indicating agreement between the model and the data. Suites of such validation tests are used as a summary of modeling progress within a scientific community. 
Building on this framework, we develop \textit{NeuronUnit}, a library that supports validation testing for neurophysiology.

@@

Today, scientists contribute a new model to the research community by submitting a paper describing how the model works, along with selected figures demonstrating its scope and validity and comparing its validity to other models with the same scope. 
Reviewers are then responsible for discovering relevant data and competing models that the paper did not adequately consider, drawing on their knowledge of prior publications. 
Unfortunately, in many areas, the number of publications being generated every year can overwhelm even the most conscientious scientists\cite{jinha_article_2010}.

@@

Professional software developers face similar issues. 
They must understand the scope of each component of a software project and validate it by measuring how well each component achieves its stated goals. 
But software developers do not validate components by simply choosing a few interesting inputs and presenting the outputs to reviewers. 
Rather, they typically follow a \emph{test-driven development} methodology by creating a suite of executable \emph{unit tests} that collectively specify the scope of each component and validate the functionality of its implementation as it is being developed and modified. 
Each test individually checks that a small portion of the program meets a single correctness criterion. For example, a unit test might verify that one function within the program correctly handles malformed inputs. 
Collectively, the test results serve as a summary of the validity of the project as it progresses through its development cycle. 
Developers can determine which features are unimplemented or buggy by examining the set of failed tests, and progress can be measured in terms of how many tests the program passes over time. 
This methodology is widely adopted in practice\cite{beck2003}.

@@

Test-driven methodologies have started to see success in science as well. 
Modeling competitions in neuroscience, for example, are typically organized around a collection of simple validation criteria, implemented as executable tests. 
These competitions continue to drive important advances and improve scientists' understanding of the relative merits of different models. 
For example, the quantitative single neuron modeling competition (QSNMC)\cite{jolivet_quantitative_2008} investigates the complexity-accuracy tradeoff among reduced models of excitable membranes; 
the ``Hopfield'' challenge\cite{hopfield_what_2000} tested techniques for generating neuronal network form given its function; 
the Neural Prediction Challenge sought the best stimulus reconstructions, given neuronal activity\cite{neural_prediction_url}; the Diadem challenge is advancing the art of neurite reconstruction\cite{diadem_url}; and 
examples from other subfields of biology abound\cite{dream_url}.
% \begin{figure}%{r}{0.6\textwidth}
% %\vspace{-38px}
% \includegraphics[scale=0.75]{sciunit_overview.pdf}
% \caption{Proposal overview. 
% \small{i) The \textit{SciUnit} framework can generate discipline-specific libraries for tests generation and model interfacing.  
% \textit{NeuronUnit} is proposed and described here.  Experimental data guides test development; model generation informs and is informed by these \textit{SciUnit} libraries.  
% Each \text{SciUnit} test suite repository on GitHub is indexed by the \textit{SciDash} web application.  
% Test suites for OSB and QSNMC (described in the text) are being built. 
% \textit{SciDash} automatically runs the test collections in each suite and publicly displays the results.}}
% %\vspace{-10px}
% \label{fig:sciunit_overview}
% \end{figure}
% \leavevmode
% \vspace{-0px}
\begin{figure}
\vspace{-38px}
\centering
\includegraphics[scale=0.6]{diagram1.pdf}
\vspace{-55px}
\caption{NeuronUnit overview. NeuronUnit is set of testing tools built upon the discipline-agnostic SciUnit framework. NeuronUnit can in principle test arbitrary neurophysiology models using arbitrary data but we provide here an example using models described in NeuroML as part of the Open Source Brain (OSB) project \cite{osb_url}, and single neuron electrophysiology data available as part of the NeuroElectro Project \cite{neuroelectro_url}.  Records of test results for various model/test combinations are accessible via SciDash, which indexes GitHub repositories of these records, models, and tests so they can be searched and filtered by the community.}  
\label{fig:sciunit_overview}
\end{figure}
\leavevmode
\vspace{-0px}

Each of these examples has leveraged \emph{ad hoc} infrastructure to support test generation. 
While the specific criteria used to evaluate models varies widely between disciplines in neuroscience, the underlying test-driven methodology has many common features that could be implemented once. 
Recognizing this, we have developed a discipline-agnostic framework for developing scientific validation test suites called \textit{SciUnit}\cite{sciunit_url}. 
To demonstrate the power of this framework and show how it can be used to bridge several existing neuroinformatics tools, we then developed \textit{NeuronUnit\-}. 
This library allows scientists to build \textit{SciUnit} tests that validate \text{NeuroML} models against electrophysiological data collected by the \textit{NeuroElectro Project} \cite{neuroelectro_url}, leveraging facilities from the \textit{NeuroTools} library to extract relevant features of model output.  This is summarized in Figure \ref{fig:sciunit_overview}, which shows the relationships between the layers described here.
% In addition to this open testing toolkit we also provide a library of tests for models of spiking single neurons.

@@

A simple \emph{family} of models, \verbx{LinearModel}, implementing the capability in Fig. \ref{fig:capability} as well as another capability shown in Fig. \ref{fig:training}, is shown in Fig. \ref{fig:simple_model}. 
Models in this family, \verbx{LinearModel}, produce a spike count by applying a linear transformation to the mean of the provided input current. 
The family is parameterized by a scale factor and offset, both scalars. 
To create a \emph{particular} model, the modeler may choose particular parameter values, just as with test families:
\begin{python}
CA1_linear_model_heuristic = LinearModel(3.0, 1.0)
\end{python}
Here, the parameters to the model were picked by the modeler heuristically, or based on externally-available knowledge. 
An alternative test design to the one in Fig. \ref{fig:rate_test} would add a second required capability (Fig. \ref{fig:training}) and a corresponding training phase to the \verb|run_test| method. 
This design would be relevant only for those models for which parameters can be adjusted without human involvement. 
Whether to build a training phase into the test protocol is a choice left to each modeler. 
Although Fig. \ref{fig:rate_test} does not include a training phase, if training data is available then models that nevertheless implement a training capability (like \verb|LinearModel|) can simply be trained explicitly by calling the capability method just like any other Python method:
\begin{python}
CA1_linear_model_fit = LinearModel()
CA1_linear_model_fit.train_with_currents(CA1_training_in, CA1_training_out)
\end{python}
\begin{figure}
\begin{python}
class LinearModel(sciunit.Model, SpikeCountFromCurrent, 
    TrainSpikeCountFromCurrent):
  def __init__(self, scale=None, offset=None): 
    self.scale, self.offset = scale, offset

@@

% Section 4.
\section{Discussion}

@@

\subsection{Arbitrary Scoring Criteria for Tests}
A test first assesses goodness-of-fit, and applies a normalization (e.g. pass/fail, 0.0-1.0) to generate a score. 
Arbitrary choices at both stages may benefit some models over others.  \textbf{First}, however, rank-ordering is constant across many goodness-of-fit metrics, meaning that choice of metric will rarely cause an otherwise passing model to fail and vice versa. 
For example, given a data mean and variance, ordering model output by Z-score or p-value will yield the same relative ranking of models. 
Indeed, rank ordering of models may prove more valuable than test scores themselves. 
\textbf{Second}, suite repositories are open (e.g. Fig. \ref{fig:scidash_repo}), so tests can be cloned and new statistical choices implemented. 
Statistics as applied to neuroscience have been notoriously ``fast and loose''; 
identification and correction of flawed methodology is becoming increasingly common\cite{button_power_2013,kriegeskorte_circular_2009,galbraith_study_2010,fish_fmri_url}, and is accelerated by an open framework. 
The validation criteria is made explicit in the specification of a test, so modelers need not guess which criteria are being used to validate or invalidate their model. 
Validation criteria are subject to debate (indeed, the QSNMC criteria changed between 2007 and 2008 due to such debates), and neuroscientists who wish to promote different criteria need only derive alternative tests. 
A community consensus will slowly emerge in the form a commonly-accepted test suite (see Sec. \ref{sec:scidash}). 
This community can judge which test version is most appropriate, i.e. what a model \textit{should} do -- this process documented via standard moderation techiniques used on GitHub -- and the \textit{SciUnit} framework determines whether the model \textit{does} it.
